Framing Effects in the Wild

### 2022-07-21 paper/Manuscript_Framing.Rmd
Update abstract

---
title: "Framing Effects in the Wild: How the News Media Shapes Public Opinion by Defining Political Issues"
author: "Nicolai Berk"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    includes:
      in_header: preamble.tex
subtitle: Manuscript prepared for the 12th Annual Conference of the European Political Science Association in Prague, Czech Republic
fontsize: 12pt
geometry: margin=1.5in
abstract: "A vast body of research is devoted to the persuasive effects of news framing. However, the study of the effects of framing has almost exclusively taken an experimental approach, with limited external validity. I provide the first observational estimate of framing effects by exploiting a unique case from Germany, where an editorial change in Germany's largest tabloid - *Bild* - led to a substantial shift in the paper's migration framing. Combining panel data from over 15,000 individuals with content analysis of 2.5 million newspaper articles using transformer models and a difference-in-differences design, I show that a substantial increase in the emphasis of crime in migration coverage did not affect immigration attitudes, nor did it increase the association of migration attitudes with attitudes about crime. These findings fundamentally challenge the generalizability of framing effects research and question worries about the news media|s power to shape public opinion."
bibliography: C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib
---

@@

WORD COUNT: 6,596

*I am thankful for extensive feedback on earlier stages of this draft from Christian Arnold, Daniel Bischof, Philipp Broniecki, Jóhanna Ýr Bjarnadóttir, Elias Dinas, Florian Foos, Felix Hagemann, Julian Heid, Christina Gahn, Heike Klüver, Martin Vinæs Larsen, Matthew Levendusky, Will Lowe, Thomas M. Meyer, Rune Slothuus, Judith Spirig, Antonio Valentim as well as participants in the EPSIP Colloquium at HU Berlin, CIVICA EGN Workshop at the EUI, CompText Conference 2022, and the Political Behaviour Section at Aarhus University. I also thank the two RAs, Robin Groß and Nathalie Aerts, for annotating the news articles. All remaining confusions are my own.*

@@

# Framing Effects vs. Media Effects

@@

@@

@@

single_model <- 
  fixest::feglm(dv ~ post*treat | lfdn + wave_ctrl, 
                data = 
                  # add control for different post-waves
                  gles_p_long %>% 
                  filter(!is.na(dv)) %>% 
                  mutate(
                    wave_ctrl = 
                      ifelse(
                        wave %in% unique(.$wave[c(1,2)]),
                        "reference",
                        wave
                      ) 
                  ) %>% 
                  select(dv, post, treat, lfdn, wave_ctrl))

@@

Figure \@ref(fig:singledv) visualizes the effect of the change in framing on immigration attitudes. Green, dashed vertical lines indicate the expected effect from the experimental literature. The orange dotted lines indicate the equivalence boundaries. If the 90% interval is contained within these boundaries, the hypothesis that the estimate is larger than the SESOI can be rejected. The point estimate indicates a `r round(ATT, 3)`-point change among Bild readers following treatment (\$t = \$ `r round(t_ATT, 2)`, \$p = \$ `r round(p_ATT, 2)`), which is around `r round(ATT*100/0.6)`% of the expected effect of 0.6. The 95% confidence intervals also contains values far smaller than the SESOI of 0.15 but clearly include the vertical red line indicating no change. Clearly, the model suggests a null effect.

```{r singleplot, fig.cap="Standardised Effect of Change in Framing on Alternative Dependent Variables", fig.height=2}

EffectPlot(multi = F, scaled = "raw", theoretical_effect_size = 0.6, boundary_share = 0.25)

EffectPlot(multi = T, scaled = T, theoretical_effect_size = 0.2) # 0.2 SD suggested by Peyton 2020, 599

@@

EffectByWavePlot(theoretical_effect_size = 0.6, boundary_share = 0.25)
# would add 2 (or 3) lines here: 
#   minimal relevant effect (5%/0.3 on 7-point scale); 
#   conservative experimental effect (10%/0.6)
#   large (not maximum) experimental effect (30%/2.4)

@@

>>>>

---
title: "Framing Effects in the Wild"
author: "Nicolai Berk"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    includes:
      in_header: preamble.tex
# subtitle: Manuscript prepared for the 12th Annual Conference of the European Political Science Association in Prague, Czech Republic
fontsize: 12pt
geometry: margin=1.5in
abstract: "A large body of evidence from experimental work suggests that individual attitudes can easily be altered by emphasizing different aspect of an the issue at stake. However, such experimental studies of *emphasis framing* are limited in their emulation of the real world, where consumers are treated by multiple sources, select their news content, and pay more or less attention compared to the experimental setting. I exploit a rare change in news framing in the major German tabloid paper - *Bild* - to study whether experimental estimates of framing effects translate to the real world. In a first step, I apply supervised transformer models to a novel dataset of 2.5M news articles and provide evidence of a substantial shift in *Bild*'s migration framing following an editorial change. Second, I use a difference-in-differences design on a panel datset of over 16,000 individuals in ten waves to assess the precise impact on migration attitudes. The estimated effect is thirty times smaller than similar estimates from experimental studies, statisticially indistinguishable from zero, and significantly smaller than even minimal relevant effect magnitudes. Further analyses show that no change can be observed on other, related dependent variables either, and suggest that the change did not result in outward or inward selection into the newspaper. Taken together, these findings fundamentally challenge the external validity of framing effects research and suggests more limited influence of the news media on public opinion than previously thought."
bibliography: C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib
---

@@

WORD COUNT:

*I am thankful for extensive feedback and other support on this manuscript from Christian Arnold, Daniel Bischof, Philipp Broniecki, Jóhanna Ýr Bjarnadóttir, Elias Dinas, Florian Foos, Felix Hagemann, Violetta Haas, Julian Heid, Christina Gahn, Heike Klüver, Martin Vinæs Larsen, Matthew Levendusky, Will Lowe, Thomas M. Meyer, Rune Slothuus, Judith Spirig, Antonio Valentim as well as participants in the EPSIP Colloquium at HU Berlin, CIVICA EGN Workshop at the EUI, CompText Conference 2022, and the Political Behaviour Section at Aarhus University. I also thank the two RAs, Robin Groß and Nathalie Aerts, for annotating the news articles. All remaining confusions are my own.*

@@

# Limits of Framing Experiments

@@

## Difference of Treated and Untreated Units [or matching]

@@

## Effect on Migration Attitude

@@

single_model <- 
  fixest::feglm(dv ~ post*treat | lfdn + wave_ctrl, 
                data = 
                  gles_p_long %>% 
                  filter(!is.na(dv)) %>% 
                  mutate(
                    wave_ctrl = 
                      ifelse(
                        wave %in% unique(.$wave[c(1,2)]),
                        "reference",
                        wave
                      ) 
                  ) %>% 
                  select(dv, post, treat, lfdn, wave_ctrl))

@@

Figure \@ref(fig:singledv) visualizes the effect of the change in framing on immigration attitudes. Dashed vertical lines indicate the expected effect from the experimental literature. The dotted lines indicate the equivalence boundaries. If the 90% interval is contained within these boundaries, the hypothesis that the estimate is larger than the SESOI can be rejected. The point estimate indicates a `r round(ATT, 3)`-point change among Bild readers following treatment (\$t = \$ `r round(t_ATT, 2)`, \$p = \$ `r round(p_ATT, 2)`), which is around `r round(ATT*100/0.6)`% of the expected effect of 0.6. The 95% confidence intervals also contains values far smaller than the SESOI of 0.15 but clearly include the vertical red line indicating no change. Clearly, the model suggests a null effect.

```{r singleplot, fig.cap="Effect of change in framing on migration attitudes; difference-in-differences estimate.", fig.height=2}

DiDPlot(multi = F, scaled = "raw", theoretical_effect_size = 0.6, boundary_share = 0.25)

@@

DiDByWavePlot(theoretical_effect_size = 0.6, boundary_share = 0.25)

## Effect on other dependent variables

DiDPlot(multi = T, scaled = T, theoretical_effect_size = 0.2) # 0.2 SD suggested by Peyton 2020, 599

@@

## Outward Selection
